### 2021-04-22 00-Data-Processing/04-Data-Analysis.Rmd
Analysis of accessibility by age is probably complete

---
title: "04-Data-Analysis"
output: html_notebook
---

Analysis of accessibility to foodbanks in Hamilton.

Load packages:
```{r load-packages, include=FALSE}
library(patchwork)
library(sf)
library(tidyverse)
library(tmap)
#library(vaccHamilton)
```

Invoke data:
```{r}
load("output-data-files/data_da_2016.RData")
load("output-data-files/dwelling_network_points_2016.RData")
load("output-data-files/hamilton_boundary.RData")
load("output-data-files/population_age_income.RData")
load("output-data-files/foodbank_locations.RData")
load("output-data-files/modes_less40k.RData")
load("output-data-files/ttm_car.RData")
load("output-data-files/ttm_transit.RData")
load("output-data-files/ttm_walk.RData")
#data("data_da_2016")
#data("dwelling_network_points_2016")
#data("hamilton_cma")
#data("modes_less40k")
#data("foodbank_locations")
#data("ttm_car")
#data("ttm_transit")
#data("ttm_walk")
```

## Remove entries in travel time matrices for inapplicable destinations

As a first step, we will remove the computed travel times to the three Community Kitchen locations (IDs 1-3):
```{r}
ttm_car <- ttm_car %>%
  filter(!between(OBJECTID, left = 1, right = 3))

ttm_transit <- ttm_transit %>%
  filter(!between(OBJECTID, left = 1, right = 3))

ttm_walk <- ttm_walk %>%
  filter(!between(OBJECTID, left = 1, right = 3))
```

## Quick check

Distribution of the number of foodbanks that can be reached by each mode at the maximum travel time used in the calculation of the time matrices (180 min or 10,000 m).

Car:
```{r}
ttm_car %>% group_by(UID) %>% summarize(fb_per_parcel = n()) %>% summary()
```

All parcels can reach every foodbank within 180 min.

Transit:
```{r}
ttm_transit %>% group_by(UID) %>% summarize(fb_per_parcel = n()) %>% summary()
```

Half of all parcels can reach 49 foodbanks or more in 180 min or less.

Walking:
```{r}
ttm_walk %>% group_by(UID) %>% summarize(fb_per_parcel = n()) %>% summary()
```

Half of all parcels can reach 42 foodbanks or more in 180 min or less.

## Prepare data

Join number of low income households per DA to dwellings and total number of residential units:
```{r}
dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
  left_join(data_da_2016 %>%
              st_drop_geometry() %>%
              dplyr::transmute(DAUID = GeoUID, 
                               p_inc_less40k),
            by = "DAUID")
```

Calculate probability that a point on the network will have low income households. This is the proportion of low income households in the DA multiplied by the total number of residential units at the point:
```{r}
dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
  mutate(income_less40k = p_inc_less40k * Dwellings)
```

Join the household data to the travel time matrices:
```{r}
ttm_car <- ttm_car %>%
  left_join(dwelling_network_points_2016 %>%
              st_drop_geometry() %>%
              dplyr::select(ID, 
                            income_less40k,
                            DAUID, 
                            TAZUID),
            by = c("UID" = "ID"))

ttm_transit <- ttm_transit %>%
  left_join(dwelling_network_points_2016 %>%
              st_drop_geometry() %>%
              dplyr::select(ID, 
                            income_less40k,
                            DAUID, 
                            TAZUID),
            by = c("UID" = "ID"))

ttm_walk <- ttm_walk %>%
  left_join(dwelling_network_points_2016 %>%
              st_drop_geometry() %>%
              dplyr::select(ID, 
                            income_less40k,
                            DAUID, 
                            TAZUID),
            by = c("UID" = "ID"))
```

Join the food bank data to the travel time matrices:
```{r}
ttm_car <- ttm_car %>%
  left_join(foodbank_locations %>%
              st_drop_geometry() %>%
              dplyr::select(ID, 
                            TYPE,
                            COVIDClose),
            by = c("OBJECTID" = "ID"))

ttm_transit <- ttm_transit %>%
  left_join(foodbank_locations %>%
              st_drop_geometry() %>%
              dplyr::select(ID, 
                            TYPE,
                            COVIDClose),
            by = c("OBJECTID" = "ID"))

ttm_walk <- ttm_walk %>%
  left_join(foodbank_locations %>%
              st_drop_geometry() %>%
              dplyr::select(ID,
                            TYPE,
                            COVIDClose),
            by = c("OBJECTID" = "ID"))
```

Drop NAs:
```{r}
ttm_car <- ttm_car %>%
  drop_na(TAZUID,
          income_less40k)

ttm_transit <- ttm_transit %>%
  drop_na(TAZUID,
          income_less40k)

ttm_walk <- ttm_walk %>%
  drop_na(TAZUID,
          income_less40k)
```

Calculate proportion of trips by mode per TAZ:
```{r}
modes <- modes_less40k %>%
  mutate(p_car = Driver/(Driver + Transit + Walk),
         p_transit = Transit/(Driver + Transit + Walk),
         p_walk = Walk/(Driver + Transit + Walk))
```

Assume that TAZ with NAs have a proportion of car trips of one:
```{r}
modes <- modes %>%
  dplyr::mutate(p_car = replace_na(p_car, 1),
                p_transit = replace_na(p_transit, 0),
                p_walk = replace_na(p_walk, 0))
```

```{r}
# join tripmaking modal proportions
ttm_car <- ttm_car %>%
  left_join(modes %>%
              st_drop_geometry() %>%
              transmute(TAZUID,
                        p_car,
                        p_transit,
                        p_walk),
            by = "TAZUID")

ttm_transit <- ttm_transit %>%
  left_join(modes %>%
              st_drop_geometry() %>%
              transmute(TAZUID,
                        p_car,
                        p_transit,
                        p_walk),
            by = "TAZUID")

ttm_walk <- ttm_walk %>%
  left_join(modes %>%
              st_drop_geometry() %>%
              transmute(TAZUID,
                        p_car,
                        p_transit,
                        p_walk),
            by = "TAZUID")
```

Prepare and join population data:
```{r number-of-residential-units-per-DA, include=FALSE}
# Data preparation
## Find number of residential units in each DA. This is needed to assign the population to parcels
residential_units <- dwelling_network_points_2016 %>%
  st_drop_geometry() %>%
  group_by(TAZUID) %>%
  summarize(residential_units = sum(Dwellings))
```

```{r join-population-to-parcels, include=FALSE}
# Data preparation
## Join number of people age 55+ per DA to dwellings and total number of residential units:
dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
  left_join(population_age_income %>%
              st_drop_geometry() %>%
              dplyr::transmute(TAZUID,
                               children,
                               adults,
                               seniors),
            by = "TAZUID") %>%
  left_join(residential_units,
            by = "TAZUID")
```

```{r calculate-population-per-parcel, include=FALSE}
# Data preparation
##Calculate population per point on the network. This is the population in the DA divided by the total number of residential units. This gives the average number of people per residential unit. Then, multiply by the number of dwellings (residential units) at the point:
dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
  mutate(children = (children/residential_units) * Dwellings,
         adults = (adults/residential_units) * Dwellings,
         seniors = (seniors/residential_units) * Dwellings)
```

```{r join-population-data-to-travel-time-tables, include=FALSE}
# Data preparation
## Join the population data to the travel time matrices:
ttm_car <- ttm_car %>%
  left_join(dwelling_network_points_2016 %>%
              st_drop_geometry() %>%
              dplyr::transmute(ID, 
                            children,
                            adults,
                            seniors),
            by = c("UID" = "ID"))

ttm_transit <- ttm_transit %>%
  left_join(dwelling_network_points_2016 %>%
              st_drop_geometry() %>%
              dplyr::transmute(ID, 
                            children,
                            adults,
                            seniors),
            by = c("UID" = "ID"))

ttm_walk <- ttm_walk %>%
  left_join(dwelling_network_points_2016 %>%
              st_drop_geometry() %>%
              dplyr::transmute(ID, 
                            children,
                            adults,
                            seniors),
            by = c("UID" = "ID"))
```

Join all tables:
```{r}
ttm <- ttm_car %>%
  rename(travel_time_car = travel_time) %>%
  full_join(ttm_transit %>%
              transmute(UID,
                        OBJECTID,
                        travel_time_transit = travel_time),
            by = c("UID", "OBJECTID")) %>%
  full_join(ttm_walk %>%
              transmute(UID,
                        OBJECTID,
                        travel_time_walk = travel_time),
            by = c("UID", "OBJECTID"))
```

Check the NAs in travel time table:
```{r}
ttm %>% 
  filter(is.na(travel_time_transit)) %>%
  select(travel_time_transit,
         p_transit) %>%
  summary()
```

What does this mean? Most DAs missing travel time by transit do not report trips by transit in the TTS. In some DAs there are trips by transit (p_transit != 0), but no foodbank can be reached by transit (travel_time_transit is NA). How many of these cases?
```{r}
ttm %>% 
  filter(is.na(travel_time_transit) & p_transit > 0) %>%
  select(travel_time_transit,
         p_transit) %>%
  summary()
```

It should be fine to replace the NAs with zeros, which multiplied by p_transit will return a zero (i.e., it is not possible to reach a foodbank by transit).
```{r}
ttm <- ttm %>% 
  mutate(travel_time_transit = replace_na(travel_time_transit, 0))
```

Check the NAs in travel time by walk:
```{r}
ttm %>% 
  filter(is.na(travel_time_walk)) %>%
  select(travel_time_walk,
         p_walk) %>%
  summary()
```

What does this mean? Most DAs missing travel time by walking do not report trips by walking in the TTS. In some DAs there are trips by walking (p_walk != 0), but no foodbank can be reached by walk (travel_time_walk is NA). How many of these cases?
```{r}
ttm %>% 
  filter(is.na(travel_time_walk) & p_walk > 0) %>%
  select(travel_time_walk,
         p_walk) %>%
  summary()
```

Like with transit, it should be fine to replace the NAs with zeros, which multiplied by p_walk will return a zero (i.e., it is not possible to reach a foodbank by walking).
```{r}
ttm <- ttm %>% 
  mutate(travel_time_walk = replace_na(travel_time_walk, 0))
```

Summary of the table after these operations:
```{r}
summary(ttm)
```

Calculate expected travel times by mode. These are the travel time by the mode multiplied by the probability of using the mode (which is proxied by the proportion of trips by mode in the DA from the TTS). The total expected travel time is the sum of the expected travel times by mode:
```{r}
ttm <- ttm %>%
  mutate(e_travel_time_car = travel_time_car * p_car,
         e_travel_time_transit = travel_time_transit * p_transit,
         e_travel_time_walk = travel_time_walk * p_walk,
         e_travel_time = e_travel_time_car + e_travel_time_transit + e_travel_time_walk)
```

What is the mean expected travel time for each parcel over all the foodbanks (open and closed)?
```{r}
e_travel_time_parcel <- ttm %>% 
  group_by(UID) %>%
  summarize(DAUID = first(DAUID),
            e_travel_time = mean(e_travel_time))

summary(e_travel_time_parcel)
```

Join geometry for plotting:
```{r}
e_travel_time_parcel <- e_travel_time_parcel %>%
  left_join(dwelling_network_points_2016 %>%
              select(ID, geometry),
            by = c("UID" = "ID")) %>%
  st_as_sf()
```

Plot:
```{r}
ggplot() +
  geom_sf(data = hamilton_boundary) +
  geom_sf(data = e_travel_time_parcel,
          aes(color = e_travel_time),
          alpha = 0.3) + 
  scale_color_distiller(palette = "OrRd", direction = 1)
```

What is the mean expected travel time for each parcel over all the foodbanks that did NOT close during COVID?
```{r}
e_travel_time_parcel_2 <- ttm %>% 
  filter(COVIDClose == "No") %>%
  group_by(UID) %>%
  summarize(DAUID = first(DAUID),
            e_travel_time = mean(e_travel_time))

summary(e_travel_time_parcel_2)
```

Join geometry for plotting:
```{r}
e_travel_time_parcel_2 <- e_travel_time_parcel_2 %>%
  left_join(dwelling_network_points_2016 %>%
              select(ID, geometry),
            by = c("UID" = "ID")) %>%
  st_as_sf()
```

Plot:
```{r}
ggplot() +
  geom_sf(data = hamilton_boundary) +
  geom_sf(data = e_travel_time_parcel_2,
          aes(color = e_travel_time),
          alpha = 0.3) + 
  scale_color_distiller(palette = "OrRd", direction = 1)
```

Notice how expected travel time to all open food banks can _decrease_ with the closures! Food banks on average are a little bit closer, but there are _fewer_ of them, which will tend to increase the demand: 
```{r}
summary(e_travel_time_parcel)
summary(e_travel_time_parcel_2)
```

## Accessibility

Define a function for calculating accessibility using a binary impedance function:
```{r}
#' Compute balanced floating catchment area accessibility
#'
#' @param ttm A travel time matrix with a column of origin IDs, a column of destination IDs, travel times, population at the origin, and supply at the destination.
#' @param threshold A threshold for the binary impedance function
#' @return The level of service \code{los} and accessibility \code{accessibility}.
#' @export
#' @examples
#' #add(1, 1)
#' #add(10, 1)

b2sfca <- function(ttm = ttm, threshold = threshold){
  # calculate the impedance
  ttm <- ttm %>%
    dplyr::mutate(impedance_binary = ifelse(travel_time <= threshold, 1, 0))

  # Calculate the sum of the impedance by population for the balancing factors. In the binary case, this should be the same as the number of SoBi hubs that each interpolated population unit can reach
  sum_b1 <- ttm %>%
    dplyr::group_by(UID) %>%
    dplyr::summarize(sum_b1 = sum(impedance_binary),
              .groups = "drop")

  # Calculate the sum of the impedance by SoBi hubs for the balancing factors. In the binary case, this should be the same as the number of interpolated population cells that each SoBi hub serves
  sum_b2 <- ttm %>%
    dplyr::group_by(OBJECTID) %>%
    dplyr::summarize(sum_b2 = sum(impedance_binary))

  # Join the sum of the impedance to the table to calculate the balanced impedance values:
  ttm <- ttm %>%
    dplyr::left_join(sum_b1, by = "UID") %>%
    dplyr::left_join(sum_b2, by = "OBJECTID")

  # **Important:** notice that some values of `sum_b1` are zeros! This means that some population units cannot reach a SoBi hub in at most 24 min. Since the balancing impedance is calculated by diving the impedance by `sum_b1`, this will lead to NaNs (divisions by zero). To avoid issues we will remove any population units that reach zero hubs:
  ttm <- ttm %>%
    dplyr::filter(sum_b1 > 0,
                  sum_b2 > 0)

  # Calculate balanced impedance values:
  ttm <- ttm %>%
    dplyr::mutate(balanced_impedance_1 = impedance_binary/sum_b1,
                  balanced_impedance_2 = impedance_binary/sum_b2)

  # The level of service of each hub is the number of bicycle racks at the hub, divided by the population that they serve:
  los <- ttm %>%
    dplyr::group_by(OBJECTID) %>%
    dplyr::summarize(los = dplyr::first(supply) / sum((population * balanced_impedance_1)),
              .groups = "drop")

  # Join the level of service to the table:
  ttm <- ttm %>%
    dplyr::left_join(los, by = "OBJECTID")

  # To calculate accessibility, the level of service of each hub needs to be distributed proportionally to the population units.
  accessibility <- ttm %>%
    dplyr::group_by(UID) %>%
    dplyr::summarize(accessibility = sum(los * balanced_impedance_2),
              .groups = "drop")

  output = list(los = los,
                accessibility = accessibility)
  return(output)
}
```

Calculate accessibility with a threshold of 15 min:
```{r}
acc_15min <- ttm %>% 
  transmute(UID, OBJECTID, travel_time = e_travel_time, population = income_less40k, supply = 1) %>%
  b2sfca(threshold = 15)
```

Check that los and accessibility balance:
```{r}
sum(acc_15min$los$los)
sum(acc_15min$accessibility$accessibility)
```

Sensitivity analysis:
```{r accessibility-sensibility}
# Testing different thresholds to check how accessibility changes
threshold_seq <- seq(from = 2, to = 30, by = 2)

accessibility_test <- data.frame(threshold = threshold_seq, 
                                 accessibility = numeric(length(threshold_seq)))
for(i in 1:length(threshold_seq)){
  results_test <-  ttm %>% 
  transmute(UID, OBJECTID, travel_time = e_travel_time, population = income_less40k, supply = 1) %>%
    b2sfca(threshold = i )
  accessibility_test$accessibility[i] <- sum(results_test$accessibility$accessibility)
}

acc_sensitivity <- ggplot(accessibility_test, 
       aes(x = threshold, y = accessibility)) +
  geom_line() +
  ylab("Accessibility")

acc_sensitivity
```

30-min accessibility pre-covid:
```{r 30-min-pre-covid, include=FALSE}
results_30min_pre <- ttm %>%
  #dplyr::filter(COVIDClose == "No") %>%
  transmute(UID, 
            OBJECTID, 
            travel_time = e_travel_time, 
            population = income_less40k, 
            supply = 1) %>%
  b2sfca(threshold = 30)
```

30-min accessibility during covid:
```{r 30-min-during-covid, include=FALSE}
results_30min_covid <- ttm %>%
  dplyr::filter(COVIDClose == "No") %>%
  transmute(UID, 
            OBJECTID, 
            travel_time = e_travel_time, 
            population = income_less40k, 
            supply = 1) %>%
  b2sfca(threshold = 30)
```

Regional accessibility pre-covid:
```{r, 30-min-sum-pre}
sum(results_30min_pre$los$los)
sum(results_30min_pre$accessibility$accessibility)
```

Regional accessibility during covid:
```{r 30-min-sum-covid}
sum(results_30min_covid$los$los)
sum(results_30min_covid$accessibility$accessibility)
```

Bind results:
```{r, bind-all-results}
los <- rbind(data.frame(results_30min_pre$los, 
                        timing = "Pre-covid", 
                        threshold = "30 min"),
             data.frame(results_30min_covid$los, 
                        timing = "During covid", 
                        threshold = "30 min")) %>%
  mutate(timing = factor(timing, 
                         levels = c("Pre-covid", 
                                    "During covid"),
                         ordered = TRUE))

accessibility <- rbind(data.frame(results_30min_pre$accessibility, 
                                  timing = "Pre-covid", 
                                  threshold = "30 min"),
                       data.frame(results_30min_covid$accessibility, 
                                  timing = "During covid", 
                                  threshold = "30 min")) %>%
  mutate(timing = factor(timing, 
                         levels = c("Pre-covid", 
                                    "During covid"),
                         ordered = TRUE))
```

Join geometry:
```{r, add-geometry-to-level-of-service, include=FALSE}
los <- los %>%
  left_join(foodbank_locations %>%
              dplyr::select(ID),
            by = c("OBJECTID" = "ID")) %>%
  st_as_sf()
```

Plot levels of service before and during covid:
```{r}
los_pre <- ggplot() +
  geom_sf(data = hamilton_boundary) +
  geom_sf(data = los %>%
            filter(timing == "Pre-covid"),
          aes(color = los,
              size = los)) +
  geom_sf(data = foodbank_locations,
          aes(shape = COVIDClose)) +
  scale_color_distiller(palette = "OrRd", 
                        direction = 1) +
  scale_shape_manual(values = c("Yes" = 4, "No" = 1))

los_covid <- ggplot() +
  geom_sf(data = hamilton_boundary) +
  geom_sf(data = los %>%
            filter(timing == "During covid"),
          aes(color = los,
              size = los)) +
  geom_sf(data = foodbank_locations,
          aes(shape = COVIDClose)) +
  scale_color_distiller(palette = "OrRd", 
                        direction = 1) +
  scale_shape_manual(values = c("Yes" = 4, "No" = 1))

los_pre / los_covid
```

Plot the change in level of service:
```{r}
los %>%
  st_drop_geometry() %>%
  pivot_wider(names_from = timing, 
              values_from = los, 
              values_fill = 0) %>%
  filter(`During covid` > 0) %>%
  mutate(los_d = `During covid` / `Pre-covid` - 1) %>%
  left_join(foodbank_locations %>%
            transmute(OBJECTID = ID,
                      geometry),
          by = "OBJECTID") %>%
  st_as_sf() %>%
  ggplot() +
  geom_sf(data = hamilton_boundary) +
  geom_sf(aes(color = los_d,
              size = los_d)) +
  geom_sf(data = foodbank_locations,
          aes(shape = COVIDClose)) +
  scale_color_distiller(palette = "OrRd", 
                        direction = -1) +
  scale_shape_manual(values = c("Yes" = 4, "No" = 1))

```

Here we can see the places where the level of service deteriorated most (excluding the places that closed, where it vanished).

Join zonal identifiers to accessibility results:
```{r}
accessibility <- accessibility %>%
  left_join(dwelling_network_points_2016 %>%
            st_drop_geometry() %>%
              dplyr::select(ID,
                            DAUID,
                            TAZUID,
                            children,
                            adults,
                            seniors),
            by = c("UID" = "ID")) %>%
  left_join(population_age_income %>%
              st_drop_geometry() %>%
              transmute(TAZUID),
            by = "TAZUID") 
```

Accessibility by TAZ:
```{r}  
accessibility_taz <- accessibility %>%
  group_by(TAZUID, 
           timing) %>%
  summarize(accessibility = sum(accessibility),
            timing = first(timing),
            .groups = "drop") %>%
  left_join(population_age_income %>%
              transmute(TAZUID,
                        children,
                        adults,
                        seniors),
            by = "TAZUID") %>%
  st_as_sf()
```

Plot accessibility before and during covid:
```{r}
acc_pre <- ggplot() +
  geom_sf(data = hamilton_boundary) +
  geom_sf(data = accessibility_taz %>%
            filter(timing == "Pre-covid"),
          aes(fill = accessibility)) +
  scale_fill_distiller(palette = "OrRd", 
                        direction = 1)

acc_covid <- ggplot() +
  geom_sf(data = hamilton_boundary) +
  geom_sf(data = accessibility_taz %>%
            filter(timing == "During covid"),
          aes(fill = accessibility)) +
  scale_fill_distiller(palette = "OrRd", 
                        direction = 1)

acc_pre / acc_covid
```

Calculate accessibility change by TAZ:
```{r}
accessibilty_taz_change <- accessibility_taz %>%
  st_drop_geometry() %>%
  pivot_wider(names_from = timing, 
              values_from = accessibility) %>%
  # Calculate changes in accessibility
  #mutate(acc_d = `During covid` / `Pre-covid` - 1) %>% # Percentage change is rather flat and not as easy to interpret as the difference
  mutate(acc_d = `During covid` - `Pre-covid`) %>%
  # Join TAZ geometry
  left_join(population_age_income %>%
            transmute(TAZUID,
                      geometry),
          by = "TAZUID") %>%
  # Convert to simple features
  st_as_sf()
```

Plot the change in accessibility by TAZ:
```{r}
  # Plot
  ggplot() +
  geom_sf(data = accessibilty_taz_change,
          aes(fill = acc_d),
          color = NA) +
  geom_sf(data = population_age_income,
          fill = NA) +
  scale_fill_fermenter(palette = "OrRd", 
                        direction = -1)
```

## Further Breakdown by Population Characteristics

Calculate accessibility change by parcel:
```{r}
accessibility_change <- accessibility %>%
  select(UID, timing, accessibility) %>%
  pivot_wider(names_from = timing, 
              values_from = accessibility,
              values_fill = 0) %>%
  # Calculate changes in accessibility
  #mutate(acc_d = `During covid` / `Pre-covid` - 1) %>% # Percentage change is rather flat and not as easy to interpret as the difference
  mutate(acc_d = `During covid` - `Pre-covid`)
```

Join accessibility change to accessibility table:
```{r}
accessibility <- accessibility %>%
  left_join(accessibility_change %>%
              select(UID, acc_d),
            by = "UID")
```

Quartiles of accessibility:
```{r}
# acc_quintiles_pre <- quantile(accessibility %>%
#                                 filter(timing == "Pre-covid") %>%
#                                 pull(accessibility),
#                               c(0, 0.25, 0.5, 0.75, 1))
# acc_quintiles_covid <- quantile(accessibility %>%
#                                 filter(timing == "During covid") %>%
#                                 pull(accessibility),
#                               c(0, 0.25, 0.5, 0.75, 1))
```

Quartiles of accessibility change:
```{r}
acc_quintiles_change <- quantile(accessibilty_change$acc_d,
                              c(0, 0.25, 0.5, 0.75, 1))
```

Add labels of accessibility levels to table:
```{r}
# accessibility <- accessibility %>%
#   mutate(accessibility_level = case_when(timing == "Pre-covid" & accessibility <= acc_quintiles_pre[2] ~ "Bottom 25%",
#                                          timing == "Pre-covid" & accessibility > acc_quintiles_pre[2] & accessibility <= acc_quintiles_pre[3] ~ "Third 25%",
#                                          timing == "Pre-covid" & accessibility > acc_quintiles_pre[3] & accessibility <= acc_quintiles_pre[4] ~ "Second 25%",
#                                          timing == "Pre-covid" & accessibility > acc_quintiles_pre[4] ~ "Top 25%",
#                                          timing == "During covid" & accessibility <= acc_quintiles_covid[2] ~ "Bottom 25%",
#                                          timing == "During covid" & accessibility > acc_quintiles_covid[2] & accessibility <= acc_quintiles_covid[3] ~ "Third 25%",
#                                          timing == "During covid" & accessibility > acc_quintiles_covid[3] & accessibility <= acc_quintiles_covid[4] ~ "Second 25%",
#                                          timing == "During covid" & accessibility > acc_quintiles_covid[4] ~ "Top 25%"),
#            accessibility_level = factor(accessibility_level,
#                                         levels = c("Top 25%", "Second 25%", "Third 25%", "Bottom 25%"),
#                                         ordered = TRUE))
```

Add labels of accessibility loss to table:
```{r}
accessibility <- accessibility %>%
  mutate(accessibility_level = case_when(acc_d <= acc_quintiles_change[2] ~ "Bottom 25%",
                                         acc_d > acc_quintiles_change[2] & acc_d <= acc_quintiles_change[3] ~ "Third 25%",
                                         acc_d > acc_quintiles_change[3] & acc_d <= acc_quintiles_change[4] ~ "Second 25%",
                                         acc_d > acc_quintiles_change[4] ~ "Top 25%",
                                         accessibility <= acc_quintiles_covid[2] ~ "Bottom 25%"),
           accessibility_level = factor(accessibility_level,
                                        levels = c("Top 25%", "Second 25%", "Third 25%", "Bottom 25%"),
                                        ordered = TRUE))
```

Calculate accessibility by age group and accessibility level:
```{r}
accessibility_by_age <- accessibility %>%
  group_by(timing, 
           accessibility_level) %>%
  summarize(children = sum(children),
            adults = sum(adults),
            seniors = sum(seniors),
            accessibility = sum(accessibility),
            .groups = "drop")

accessibility_by_age
```

The accessibility level is the magnitude of the accessibility loss: the top 25% faced the smallest losses in accessibility, and the bottom 25% faced the greatest losses in accessibility.

```{r}
cbind(accessibility_by_age %>% 
        filter(timing == "Pre-covid") %>% 
        select(children:accessibility) %>% 
        rename(accessibility_pre = accessibility),
      accessibility_by_age %>% 
        filter(timing == "During covid") %>% 
        transmute(accessibility_covid = accessibility)) %>%
  mutate(change = accessibility_covid - accessibility_pre)
```

Sanity check:
```{r}
accessibility_by_age %>%
  group_by(timing) %>%
  summarize(children = sum(children),
            adults = sum(adults),
            seniors = sum(seniors),
            accessibility = sum(accessibility))
```

### Pre-COVID Scenario to Meal Services

Get weighted minimum travel times in the pre-covid scenario:
```{r minimum-travel-time-baseline, include=FALSE}
# retrieve minimum travel time for pre-covid locations
min_ttm_car_baseline <- ttm_car %>%
  #dplyr::filter(COVIDClose == "No") %>%
  filter(TYPE != "Food Banks") %>%
  group_by(UID) %>%
  summarize(travel_time = min(travel_time),
            DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            p_car = first(p_car),
            p_transit = first(p_transit),
            p_walk = first(p_walk),
            children = first(children),
            adults = first(adults),
            seniors = first(seniors),
            .groups = "drop") %>%

  ## Calculate weighted travel time:
  mutate(weighted_travel_time = travel_time * p_car)

min_ttm_transit_baseline <- ttm_transit %>%
  #dplyr::filter(COVIDClose == "No") %>%
  filter(TYPE != "Food Banks") %>%
  group_by(UID) %>%
  summarize(travel_time = min(travel_time),
            DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            p_car = first(p_car),
            p_transit = first(p_transit),
            p_walk = first(p_walk),
            children = first(children),
            adults = first(adults),
            seniors = first(seniors),
            .groups = "drop") %>%

  ## Calculate weighted travel time:
  mutate(weighted_travel_time = travel_time * p_transit)

min_ttm_walk_baseline <- ttm_walk %>%
  #dplyr::filter(COVIDClose == "No") %>%
  filter(TYPE != "Food Banks") %>%
  group_by(UID) %>%
  summarize(travel_time = min(travel_time),
            DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            p_car = first(p_car),
            p_transit = first(p_transit),
            p_walk = first(p_walk),
            children = first(children),
            adults = first(adults),
            seniors = first(seniors),
            .groups = "drop") %>%

  ## Calculate weighted travel time:
  mutate(weighted_travel_time = travel_time * p_walk)
```

Bind the tables and label by mode:
```{r weighted-travel-time-by-mode-baseline, include=FALSE}
# Baseline analysis (pre-COVID)
## Bind the three tables and label them by mode:
min_ttm_mode_baseline <- rbind(data.frame(min_ttm_car_baseline, Mode = "Car"),
                               data.frame(min_ttm_transit_baseline, Mode = "Transit"),
                               data.frame(min_ttm_walk_baseline, Mode = "Walking"))
```

Calculate the weighted person hours of travel by age group:
```{r aggregate-weighted-travel-times-baseline, include=FALSE}
# Baseline analysis (pre-COVID)
## Aggregate weighted travel times by mode:
min_ttm_mode_baseline <- min_ttm_mode_baseline %>%
  group_by(UID) %>%
  summarize(DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            children = first(children),
            adults = first(adults),
            seniors = first(seniors),
            # Weighted travel time is the sum of the travel times weighted by the proportion of trips by mode
            weighted_travel_time = sum(weighted_travel_time)) %>%

  # add geometry
  #left_join(population_age_income %>%
  #            transmute(geometry, TAZUID),
  #          by = "TAZUID") %>%

  # calculate person-weighted hours
  mutate(wph_children = children*weighted_travel_time/60,
         wph_adults = adults*weighted_travel_time/60,
         wph_seniors = seniors*weighted_travel_time/60)
```

Calculate the weighted person hours of travel by TAZ by population group:
```{r summarize-results-by-taz-baseline, include=FALSE}
# Baseline analysis (pre-COVID)
## Summarize by TAZ and join geometry:
min_ttm_baseline_taz <- min_ttm_mode_baseline %>%
  group_by(TAZUID) %>%
  summarize(weighted_travel_time = mean(weighted_travel_time),
            wph_children = sum(wph_children),
            wph_adults = sum(wph_adults),
            wph_seniors = sum(wph_seniors)) %>%

  left_join(population_age_income %>%
              transmute(geometry, TAZUID),
            by = "TAZUID") %>%

  st_as_sf()
```

Plot weighted travel time to nearest facility:
```{r weighted-travel-time-map-baseline}
# Baseline analysis (pre-COVID)
## Map weighted travel time:
ttm_map_baseline <- ggplot() +
  geom_sf(data = min_ttm_baseline_taz,
          aes(fill = weighted_travel_time),
          color = NA) +
  scale_fill_fermenter(name = "Expected travel time (min)",
                       palette = "OrRd", 
                       direction = 1)

ttm_map_baseline
```

Plot the weighted person hours of travel for the population groups in low income households (travel times in minutes):
```{r}
tm_shape(min_ttm_baseline_taz) +
  tm_fill("wph_children")

tm_shape(min_ttm_baseline_taz) +
  tm_fill("wph_adults")

tm_shape(min_ttm_baseline_taz) +
  tm_fill("wph_seniors")
```

### Do it again for the during-COVID scenario:

Calculate weighted travel time to the nearest facilty by mode:
```{r minimum-travel-time-covid, include=FALSE}
# retrieve minimum travel time during covid to sites that didn't close
min_ttm_car_covid <- ttm_car %>%
  dplyr::filter(COVIDClose == "No", TYPE != "Food Banks") %>%
  group_by(UID) %>%
  summarize(travel_time = min(travel_time),
            DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            p_car = first(p_car),
            p_transit = first(p_transit),
            p_walk = first(p_walk),
            children = first(children),
            adults = first(adults),
            seniors = first(seniors),
            .groups = "drop") %>%

  ## Calculate weighted travel time:
  mutate(weighted_travel_time = travel_time * p_car)

min_ttm_transit_covid <- ttm_transit %>%
  dplyr::filter(COVIDClose == "No", TYPE != "Food Banks") %>%
  group_by(UID) %>%
  summarize(travel_time = min(travel_time),
            DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            p_car = first(p_car),
            p_transit = first(p_transit),
            p_walk = first(p_walk),
            children = first(children),
            adults = first(adults),
            seniors = first(seniors),
            .groups = "drop") %>%

  ## Calculate weighted travel time:
  mutate(weighted_travel_time = travel_time * p_transit)

min_ttm_walk_covid <- ttm_walk %>%
  dplyr::filter(COVIDClose == "No", TYPE != "Food Banks") %>%
  group_by(UID) %>%
  summarize(travel_time = min(travel_time),
            DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            p_car = first(p_car),
            p_transit = first(p_transit),
            p_walk = first(p_walk),
            children = first(children),
            adults = first(adults),
            seniors = first(seniors),
            .groups = "drop") %>%

  ## Calculate weighted travel time:
  mutate(weighted_travel_time = travel_time * p_walk)
```

Bind the tables and label by mode:
```{r weighted-travel-time-by-mode-covid, include=FALSE}
# Baseline analysis (pre-COVID)
## Bind the three tables and label them by mode:
min_ttm_mode_covid <- rbind(data.frame(min_ttm_car_covid, Mode = "Car"),
                               data.frame(min_ttm_transit_covid, Mode = "Transit"),
                               data.frame(min_ttm_walk_covid, Mode = "Walking"))
```

Calculate the weighted person hours of travel by age group:
```{r aggregate-weighted-travel-times-covid, include=FALSE}
# Baseline analysis (pre-COVID)
## Aggregate weighted travel times by mode:
min_ttm_mode_covid <- min_ttm_mode_covid %>%
  group_by(UID) %>%
  summarize(DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            children = first(children),
            adults = first(adults),
            seniors = first(seniors),
            weighted_travel_time = sum(weighted_travel_time)) %>%

  # add geometry
  #left_join(population_age_income %>%
  #            transmute(geometry, TAZUID),
  #          by = "TAZUID") %>%

  # calculate person-weighted hours
  mutate(wph_children = children*weighted_travel_time/60,
         wph_adults = adults*weighted_travel_time/60,
         wph_seniors = seniors*weighted_travel_time/60)
```

Aggregate the number of weighted person hours of travel by age for each TAZ:
```{r summarize-results-by-taz-covid, include=FALSE}
# Baseline analysis (PILOT LOCATIONS ONLY)
## Summarize by TAZ and join geometry:
min_ttm_covid_taz <- min_ttm_mode_covid %>%
  group_by(TAZUID) %>%
  summarize(weighted_travel_time = mean(weighted_travel_time),
            wph_children = sum(wph_children),
            wph_adults = sum(wph_adults),
            wph_seniors = sum(wph_seniors)) %>%

  left_join(population_age_income %>%
              transmute(geometry, TAZUID),
            by = "TAZUID") %>%

  st_as_sf()
```

Map weighted travel time to nearest facility:
```{r weighted-travel-time-map-covid}
# Baseline analysis (PILOT LOCATIONS ONLY)
## Map weighted travel time:
ttm_map_covid <- ggplot() +
  geom_sf(data = min_ttm_covid_taz,
          aes(fill = weighted_travel_time),
          color = NA) +
  scale_fill_distiller(name = "Expected travel time (min)",
                       palette = "OrRd", 
                       direction = 1)

ttm_map_covid
```

```{r}
tm_shape(min_ttm_covid_taz) +
  tm_fill("wph_children")

tm_shape(min_ttm_covid_taz) +
  tm_fill("wph_adults")

tm_shape(min_ttm_covid_taz) +
  tm_fill("wph_seniors")
```

### Map Differences

Join the min travel time tables together:
```{r}
min_ttm_d <- min_ttm_baseline_taz %>% 
  left_join(min_ttm_covid_taz %>%
              st_drop_geometry() %>%
              transmute(TAZUID,
                        wph_children_covid = wph_children,
                        wph_adults_covid = wph_adults,
                        wph_seniors_covid = wph_seniors), 
            by = "TAZUID") %>%
  mutate(wph_children_d = wph_children_covid / wph_children - 1,
         wph_children_d = replace_na(wph_children_d, 0),
         wph_adults_d = wph_adults_covid / wph_adults - 1,
         wph_adults_d = replace_na(wph_adults_d, 0),
         wph_seniors_d = wph_seniors_covid / wph_seniors - 1,
         wph_seniors_d = replace_na(wph_seniors_d, 0))

tm_shape(min_ttm_d) + 
  tm_fill("wph_children_d")

tm_shape(min_ttm_d) + 
  tm_fill("wph_adults_d")

tm_shape(min_ttm_d) + 
  tm_fill("wph_seniors_d")
```