### 2021-04-24 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton-title-page.tex

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{Social Science and Medicine} % Sets Journal name

\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Changes in accessibility to food banks and food services during COVID-19 and implications for low income populations in Hamilton, Ontario},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}

% Pandoc citation processing

% Pandoc header
\usepackage[table]{xcolor}

\begin{document}
\begin{frontmatter}

  \title{Changes in accessibility to food banks and food services during
COVID-19 and implications for low income populations in Hamilton,
Ontario}
    \author[University of Toronto Scarborough]{Christopher D. Higgins}
   \ead{cd.higgins@utoronto.ca} 
    \author[McMaster University]{Antonio Paez\corref{1}}
   \ead{paezha@mcmaster.ca} 
    \author[University of Toronto Mississauga]{Gyoorie Kim}
   \ead{gyoorie.kim@mail.utoronto.ca} 
    \author[University of Toronto Mississauga]{Jue Wang}
   \ead{gis.wang@utoronto.ca} 
      \address[University of Toronto Scarborough]{Department of
Geography \& Planning, University of Toronto Scarborough, 1265 Military
Trail, Toronto, ON M1C1A4}
    \address[McMaster University]{School of Earth, Environment and
Society, McMaster University, Hamilton, ON, L8S 4K1, Canada}
    \address[University of Toronto Mississauga]{Department of Geography,
Geomatics and Environment, University of Toronto Mississauga, 3359
Mississauga Road, Mississauga, ON, L5L 1C6 , Canada}
      \cortext[1]{Corresponding Author}

  \begin{abstract}
  In this paper we analyze the changes in accessibility to food banks
  and related services before and during the COVID-19 pandemic in the
  City of Hamilton, Ontario. Food banks and services are the last line
  of support for households facing food insecurity; as such, they have
  become essential and their relevance cannot be ignored in the midst of
  the economic upheaval caused by the pandemic. Our analysis is based on
  the application of balanced floating catchment areas, and concentrates
  on households with lower incomes (\$\textless CAD40,000, approximately
  the Low Income Cutoff Value for a city of Hamilton's size).
  Accessibility was low to begin with in suburban and exurban parts of
  the city. We find that about 14\% of locations originally available in
  Hamilton closed during the pandemic; the impact of those closures on
  the level of service and accessibility was disproportionate, with
  system-wide losses exceeding 40\%. Those losses were geographically
  and demographically uneven. While every part of the city faced a
  reduction in accessibility, inner suburbs fared worse in terms of loss
  of accessibility. As well, children (age \(\le 18\)) appear to have
  been impacted the most.
  \end{abstract}

 \end{frontmatter}

\hypertarget{credit-author-statement}{%
\section{CRediT author statement}\label{credit-author-statement}}

\textbf{Author 1:} Conceptualization, Methodology, Software, Validation,
Formal analysis, Investigation, Data Curation, Writing - Original Draft,
Visualization, Supervision; \textbf{Author 2:} Conceptualization,
Methodology, Software, Validation, Formal analysis, Investigation, Data
Curation, Writing - Original Draft, Visualization; \textbf{Author 3:}
Conceptualization, Investigation, Data Curation, Writing - Original
Draft; \textbf{Author 4:} Resources, Writing - Original Draft,
Supervision

\end{document}