### 2021-08-29 Accessibility-Foodbanks-Hamilton/CRediT-statement.Rmd

---
title: "CRediT Statement"
author: ""
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
bibliography: references.bib
header-includes:
  - \usepackage{changepage}
---

<!--
Conceptualization $\cdot$ Methodology	$\cdot$ Software $\cdot$ Validation	$\cdot$ Formal analysis	$\cdot$ Investigation	$\cdot$ Resources	$\cdot$ Data Curation	$\cdot$ Writing - Original Draft $\cdot$ Writing - Review & Editing	$\cdot$ Visualization $\cdot$ Supervision $\cdot$ Project administration $\cdot$ Funding 
-->

Christopher D. Higgins: Conceptualization $\cdot$ Methodology	$\cdot$ Software $\cdot$ Validation	$\cdot$ Formal analysis	$\cdot$ Investigation	$\cdot$ Data Curation	$\cdot$ Writing - Original Draft $\cdot$ Visualization $\cdot$ Supervision

Antonio Páez: Conceptualization $\cdot$ Methodology	$\cdot$ Software $\cdot$ Validation	$\cdot$ Formal analysis	$\cdot$ Investigation	$\cdot$ Data Curation	$\cdot$ Writing - Original Draft $\cdot$ Visualization

Gyoorie Kim: Conceptualization $\cdot$ Investigation	$\cdot$ Data Curation	$\cdot$ Writing - Review & Editing

Jue Wang: Conceptualization $\cdot$ Resources	$\cdot$ Writing - Review & Editing	$\cdot$ Supervision